---
title: "Problem Set 6"
author: "Subeom Lee"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    df_print: paged
linkcolor: cyan
fontsize: 10pt
colorlinks: yes
---

```{r, message=FALSE, echo=FALSE}
library(lubridate)
library(texreg)
library(broom)
library(knitr)
library(ggpubr)
library(ggrepel)
library(janitor)
opts_chunk$set(echo=TRUE,
               cache.lazy=FALSE,
               collapse=TRUE,
               comment=NA,
               warning=FALSE,
               prompt=FALSE,
               message=FALSE
               ,cache=TRUE
               )
options(digits=4, scipen=0)
masterdoc <- knitr::current_input()
```


# Questions

\footnotesize
\medskip
```{r}
#http://asbcllc.com/nbastatR/index.html

library(nbastatR)
library(future)
library(stringi)
library(tidyverse)
library(lubridate)
library(texreg)
library(broom)
library(knitr)
library(ggpubr)
library(ggrepel)
library(janitor)
library(plotly)

plan(multiprocess) 

# Run only when needed
# game_logs(seasons = 1947:2019, result_types = c("team", "player"))
# dataGameLogsTeam$Team = substring(dataGameLogsTeam$slugMatchup, 1, 3)

# Run when you updated data
# save(df_nba_player_dict, file='df_nba_player_dict.Rdata')
# save(dataGameLogsTeam, file='dataGameLogsTeam.Rdata')
# save(dataGameLogsPlayer, file='dataGameLogsPlayer.Rdata')

load('df_nba_player_dict.Rdata')
load('dataGameLogsTeam.Rdata')
load('dataGameLogsPlayer.Rdata')




avg <- aggregate(dataGameLogsTeam[, 24:46], list(dataGameLogsTeam$yearSeason, dataGameLogsTeam$Team), mean)
colnames(avg)[1] <- "Year"
colnames(avg)[2] <- "Team"



avgplot <- avg %>%
            filter(Team %in% c('GSW', 'CHI', 'HOU', 'LAL')) %>%
            ggplot(aes(x=Year, y=pctFG3Team, colour=Team)) +
            geom_line()
avgplot

avg2 <- aggregate(dataGameLogsTeam[, 24:46], list(dataGameLogsTeam$yearSeason), mean)
colnames(avg2)[1] <- "Year"
min2 <- aggregate(dataGameLogsTeam[, 24:46], list(dataGameLogsTeam$yearSeason), min)
colnames(min2)[1] <- "Year"
max2 <- aggregate(dataGameLogsTeam[, 24:46], list(dataGameLogsTeam$yearSeason), max)
colnames(max2)[1] <- "Year"

avgplot2 <- avg2 %>%
              filter(Year >= 1986) %>%
              ggplot(aes(x=Year, y=pctFG3Team*100)) +
              geom_path(colour='violet', size=2)
avgplot2

avgplot3 <- avg2 %>%
              ggplot(aes(x=Year, y=pctFG2Team*100)) +
              geom_path(colour='red', size=2)
avgplot3

avgminmax <- ggplot() +
                geom_line(data=min2, aes(x=Year, y=pctFG3Team*100), size=1, colour='green') +
                geom_line(data=max2, aes(x=Year, y=pctFG2Team*100), size=1, colour='red') +
                geom_smooth(data=avg2, aes(x=Year, y=pctFG3Team*100), size=2, colour='black')
avgminmax

avg1986 <- avg %>% filter(Year>=1986)
avg21986 <- avg2 %>% filter(Year>=1986)

avgcombinedall <- ggplot() +
                geom_line(data=avg1986, aes(x=Year, y=pctFG3Team*100, colour=Team), size=0.5, show.legend=FALSE) +
                geom_line(data=avg1986, aes(x=Year, y=pctFG2Team*100, colour=Team), size=0.5, show.legend=FALSE) +
                geom_line(data=avg21986, aes(x=Year, y=pctFG3Team*100), size=2, colour='black')+
                geom_line(data=avg21986, aes(x=Year, y=pctFG2Team*100), size=2, colour='black')
avgcombinedall

avgfiltered <- avg %>% filter(Team %in% c('GSW', 'CHI', 'HOU', 'LAL'))

avgcombined <- ggplot() +
                geom_line(data=avgfiltered, aes(x=Year, y=pctFG3Team*100, colour=Team), size=1, show.legend=FALSE) +
                geom_line(data=avgfiltered, aes(x=Year, y=pctFG2Team*100, colour=Team), size=1, show.legend=FALSE) +
                geom_smooth(data=avg2, aes(x=Year, y=pctFG3Team*100), size=2, colour='black')
avgcombined

avgcombined2 <- avgcombined +
                geom_smooth(data=avg2, aes(x=Year, y=pctFG2Team*100), size=2, colour='red')
avgcombined2

avgcombined3 <- avgcombined2 +
                geom_smooth(data=avg2, aes(x=Year, y=pctFTTeam*100), size=2, colour='green')
avgcombined3

avgcombined4 <- avgcombined3 +
                geom_smooth(data=avg2, aes(x=Year, y=ptsTeam), size=2, colour='purple')
avgcombined4

# ggplotly(p=ggplot2::last_plot())

# library(ggplot2)
# library(ggpubr)
# theme_set(theme_pubr())
# 
# figure <- ggarrange(avgplot, avgplot2,
#                     labels = c("Each Team", "All Teams"),
#                     ncol = 1, nrow = 2)
# figure

# climate <- read.csv('ps5_data.csv')
# a <- ggplot(climate) +
#       xlab('Year') +
#       ylab('Temperature(°C)') + 
#       theme(panel.border=element_rect(colour="black", fill=NA), panel.background=element_rect(fill=NA),
#             panel.grid=element_line(color="grey")) + 
#       geom_smooth(aes(Year, Lowess.5.), colour="blue", size=1) +
#       geom_line(aes(Year, No_Smoothing), colour="grey", size=1) + 
#       geom_point(aes(Year, No_Smoothing), shape=1, size=3)
# 

```
\normalsize

# Team level questions

Q1.	It seems that players are getting better at making 3-pointers than 20 years ago (both on average and also top 3-pointer shooters vs. top 3-pointer shooters) Is it true?

```{r}
fg3year <- aggregate(dataGameLogsTeam[, 35:36], list(dataGameLogsTeam$yearSeason), sum)
colnames(fg3year)[1] <- "Year"
fg3year <- fg3year %>% filter (Year >= 1986)
fg3year$pctfg3 <- fg3year$fg3mTeam / fg3year$fg3aTeam * 100

fg3yearteam <- aggregate(dataGameLogsTeam[, 35:36], list(dataGameLogsTeam$yearSeason, dataGameLogsTeam$Team), sum)
colnames(fg3yearteam)[1] <- "Year"
colnames(fg3yearteam)[2] <- "Team"
fg3yearteam <- fg3yearteam %>% filter (Year >= 1986)
fg3yearteam$pctfg3 <- fg3yearteam$fg3mTeam / fg3yearteam$fg3aTeam * 100

xaxisbreaks <- seq(1985, 2019, by=3)
yaxisbreaks <- seq(15, 45, by=5)

Q1 <- ggplot() +
  geom_line(data=fg3yearteam, aes(x=Year, y=pctfg3, colour=Team), size=0.5, show.legend=FALSE, alpha=0.7) +
  geom_line(data=fg3year, aes(x=Year, y=pctfg3), size=2, colour='black') +
  xlab('Year') +
  ylab(NULL) +
  ggtitle('3 Pointer Field Goal Success Rate') +
  theme(panel.background=element_rect(fill=NA), panel.grid.major.y=element_line(color="grey", linetype=2),
        plot.title = element_text(hjust = 0.5)) + 
  scale_y_continuous(limits=c(15, 45), breaks=yaxisbreaks, labels=paste(yaxisbreaks,"%")) +
  scale_x_continuous(limits=c(1985,2019), breaks=xaxisbreaks) +
  geom_hline(yintercept=min(fg3year$pctfg3), linetype=2, color="steelblue", size=0.5, alpha=0.9) +
  geom_hline(yintercept=max(fg3year$pctfg3), linetype=2, color="steelblue", size=0.5, alpha=0.9) +
  geom_hline(yintercept=min(fg3yearteam$pctfg3), linetype=3, color="pink", size=0.5, alpha=0.9) +
  geom_hline(yintercept=max(fg3yearteam$pctfg3), linetype=3, color="pink", size=0.5, alpha=0.9) +
  annotate("text", x=1985, y=min(fg3year$pctfg3)+0.6, label=paste(toString(round(min(fg3year$pctfg3), digits=2)),"%"), color="steelblue", size=3) +
  annotate("text", x=1985, y=max(fg3year$pctfg3)+0.6, label=paste(toString(round(max(fg3year$pctfg3), digits=2)),"%"), color="steelblue", size=3) +
  annotate("text", x=1985, y=min(fg3yearteam$pctfg3)+0.6, label=paste(toString(round(min(fg3yearteam$pctfg3), digits=2)),"%"), color="pink", size=3) +
  annotate("text", x=1985, y=max(fg3yearteam$pctfg3)+0.6, label=paste(toString(round(max(fg3yearteam$pctfg3), digits=2)),"%"), color="pink", size=3)  
  
Q1
```

Yes, the success rate of 3 point field goal has been increased by about 9% since 1986.

Q2.	If true, what could be the reasons for that?
    -	What are the expected average points of 3-pointers and 2-pointers? Show the historical data.
    -	If the expected average point from 3-pointers is getting higher than that of
      2-pointers, how should each team’s strategy changes

https://www.nytimes.com/2016/01/21/sports/basketball/how-the-nba-3-point-shot-went-from-gimmick-to-game-changer.html

Its debut, in the 1979-80 season, was inauspicious.

There are many reasons for the rise of the 3-point shot, but one may simply be math. It took a while, but coaches finally stopped listening to the traditionalist naysayers and realized that a shot that is worth 50 percent more pays off, even if that shot is a little harder to make.

“Teams have all caught on to the whole points-per-possession argument,” Lawrence Frank, the Nets’ coach at the time, said in 2009 as the 3 rate began to rapidly increase.

```{r}
fgyear <- aggregate(dataGameLogsTeam[, 35:38], list(dataGameLogsTeam$yearSeason), sum)
colnames(fgyear)[1] <- "Year"
fgyear <- fgyear %>% filter (Year >= 1986)
fgyear$pctfg3 <- fgyear$fg3mTeam / fgyear$fg3aTeam * 100
fgyear$pctfg2 <- fgyear$fg2mTeam / fgyear$fg2aTeam * 100

fgyearteam <- aggregate(dataGameLogsTeam[, 35:38], list(dataGameLogsTeam$yearSeason, dataGameLogsTeam$Team), sum)
colnames(fgyearteam)[1] <- "Year"
colnames(fgyearteam)[2] <- "Team"
fgyearteam <- fgyearteam %>% filter (Year >= 1986)
fgyearteam$pctfg3 <- fgyearteam$fg3mTeam / fgyearteam$fg3aTeam * 100
fgyearteam$pctfg2 <- fgyearteam$fg2mTeam / fgyearteam$fg2aTeam * 100

xaxisbreaks <- seq(1985, 2019, by=3)
yaxisbreaks <- seq(15, 60, by=5)

Q2_1 <- ggplot() +
  geom_line(data=fgyearteam, aes(x=Year, y=pctfg3, colour=Team), size=0.5, show.legend=FALSE, alpha=0.7) +
  geom_line(data=fgyear, aes(x=Year, y=pctfg3), size=2, colour='black') +
  geom_line(data=fgyearteam, aes(x=Year, y=pctfg2, colour=Team), size=0.5, show.legend=FALSE, alpha=0.7) +
  geom_line(data=fgyear, aes(x=Year, y=pctfg2), size=2, colour='black') +
  xlab('Year') +
  ylab(NULL) +
  ggtitle('Field Goal Success Rate') +
  theme(panel.background=element_rect(fill=NA), panel.grid.major.y=element_line(color="grey", linetype=2),
        plot.title = element_text(hjust = 0.5)) + 
  scale_y_continuous(limits=c(15, 60), breaks=yaxisbreaks, labels=yaxisbreaks) +
  scale_x_continuous(limits=c(1985,2019), breaks=xaxisbreaks)# +
  # geom_hline(yintercept=min(fg3year$pctfg3), linetype=2, color="steelblue", size=0.5, alpha=0.9) +
  # geom_hline(yintercept=max(fg3year$pctfg3), linetype=2, color="steelblue", size=0.5, alpha=0.9) +
  # geom_hline(yintercept=min(fg3yearteam$pctfg3), linetype=3, color="pink", size=0.5, alpha=0.9) +
  # geom_hline(yintercept=max(fg3yearteam$pctfg3), linetype=3, color="pink", size=0.5, alpha=0.9) +
  # annotate("text", x=1985, y=min(fg3year$pctfg3)+0.6, label=paste(toString(round(min(fg3year$pctfg3), digits=2)),"%"), color="steelblue", size=3) +
  # annotate("text", x=1985, y=max(fg3year$pctfg3)+0.6, label=paste(toString(round(max(fg3year$pctfg3), digits=2)),"%"), color="steelblue", size=3) +
  # annotate("text", x=1985, y=min(fg3yearteam$pctfg3)+0.6, label=paste(toString(round(min(fg3yearteam$pctfg3), digits=2)),"%"), color="pink", size=3) +
  # annotate("text", x=1985, y=max(fg3yearteam$pctfg3)+0.6, label=paste(toString(round(max(fg3yearteam$pctfg3), digits=2)),"%"), color="pink", size=3)  
  # 
Q2_1

```

The expected points of 2-point shots in 1986 was 'r fgyear$pctfg2[1986-1985]/100' * 2 = 'r fgyear$pctfg2[1986-1985]/100*2'
The expected points of 3-point shots in 1986 was 'r fgyear$pctfg3[1986-1985]/100' * 3 = 'r fgyear$pctfg3[1986-1985]/100*3'

The expected points of 2-point shots in 2019 was 'r fgyear$pctfg2[2019-1985]/100' * 2 = 'r fgyear$pctfg2[2019-1985]/100*2'
The expected points of 3-point shots in 2019 was 'r fgyear$pctfg3[2019-1985]/100' * 3 = 'r fgyear$pctfg3[2019-1985]/100*3'

Teams started to focus on 3-point shots after its first introduction in 1979,
because the expected points of 3-point shots are higher than that of 2-point shots
since early 90's.

```{r}
fgyear$e2 = fgyear$pctfg2 / 100 * 2
fgyear$e3 = fgyear$pctfg3 / 100 * 3

xaxisbreaks <- seq(1985, 2019, by=3)
yaxisbreaks <- seq(0.8, 1.15, by=0.05)

Q2_2 <- ggplot() +
  geom_line(data=fgyear, aes(x=Year, y=e3), size=1, colour='red') +
  geom_line(data=fgyear, aes(x=Year, y=e2), size=1, colour='blue') +
  xlab('Year') +
  ylab(NULL) +
  ggtitle('Expected points') +
  theme(panel.background=element_rect(fill=NA), panel.grid.major.y=element_line(color="grey", linetype=2),
        plot.title = element_text(hjust = 0.5)) + 
  scale_y_continuous(limits=c(0.8, 1.15), breaks=yaxisbreaks, labels=yaxisbreaks) +
  scale_x_continuous(limits=c(1985,2019), breaks=xaxisbreaks)

Q2_2

```

Q3.	Teams with more 3-pointers tend to be the better performing teams?
    -	Any insights between standings and 3-pointers?


```{r}
standings <- read_csv("standings.csv")

fgyearteam <- aggregate(dataGameLogsTeam[, 35:38], list(dataGameLogsTeam$yearSeason, dataGameLogsTeam$nameTeam), sum)
colnames(fgyearteam)[1] <- "Year"
colnames(fgyearteam)[2] <- "nameTeam"
fgyearteam <- fgyearteam %>% filter (Year >= 1986)
fgyearteam$pctfg3 <- fgyearteam$fg3mTeam / fgyearteam$fg3aTeam * 100
fgyearteam$pctfg2 <- fgyearteam$fg2mTeam / fgyearteam$fg2aTeam * 100

standings2 <- left_join(standings, fgyearteam, by=c("Year" = "Year", "Team" = "nameTeam"))

Q3 <- ggplot(standings2) +
  geom_point(aes(x=pctfg3, y=Rk), color="steelblue") +
  geom_point(aes(x=pctfg2, y=Rk), color="pink")
  # geom_line(data=fgyear, aes(x=Year, y=e2), size=1, colour='blue') +
  # xlab('Year') +
  # ylab(NULL) +
  # ggtitle('Expected points') +
  # theme(panel.background=element_rect(fill=NA), panel.grid.major.y=element_line(color="grey", linetype=2),
  #       plot.title = element_text(hjust = 0.5)) + 
  # scale_y_continuous(limits=c(0.8, 1.15), breaks=yaxisbreaks, labels=yaxisbreaks) +
  # scale_x_continuous(limits=c(1985,2019), breaks=xaxisbreaks)
  
Q3

linearModel <- lm(Rk ~ pctfg3, data=standings2)
summary(linearModel)

linearModel2 <- lm(Rk ~ pctfg2, data=standings2) 
summary(linearModel2)

linearModel3 <- lm(Rk ~ pctfg3 + pctfg2, data=standings2)
summary(linearModel3)

plot(linearModel)
plot(linearModel2)
plot(linearModel3)

linearModel4 <- lm(pctfg3 ~ pctfg2, data=standings2)
summary(linearModel4)
plot(linearModel4)

Q3_2 <- ggplot(standings2) +
  geom_point(aes(x=pctfg2, y=pctfg3), color="steelblue")
Q3_2
 
```

Yes. However, pctfg2 is more relevant than pctfg3



-	Focus on three point shooting is a strategy that started fairly recently, we can create a map to show where this strategy initially emerged and how fast it spreaded across the entire country.


```{r}
```

# Player level questions

-.	Players who are good at 3-pointers are also good at 2-pointers or free throws?

-. Are there any relationship between players’ ages and 3-pointers? Both total and average.

-	Players with high salaries are good at 3-pointers?

-	We want to analyze whether players can drastically improve their three point shooting skills over time or the skill is rather something people are borned with.

-	Show the 3-pointer statistics geographically based on players’ hometowns. Maybe this help illustrates the different basketball playing style across different regions, both domestic and international.

-	We would like to explore the importance of three point shooters in a given team by measuring the share of the team’s total salary over time.

