---
title: "Problem Set 6"
author: "Subeom Lee"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    df_print: paged
linkcolor: cyan
fontsize: 10pt
colorlinks: yes
---

```{r, message=FALSE, echo=FALSE}
library(lubridate)
library(texreg)
library(broom)
library(knitr)
library(ggpubr)
library(ggrepel)
library(janitor)
opts_chunk$set(echo=TRUE,
               cache.lazy=FALSE,
               collapse=TRUE,
               comment=NA,
               warning=FALSE,
               prompt=FALSE,
               message=FALSE
               ,cache=TRUE
               )
options(digits=4, scipen=0)
masterdoc <- knitr::current_input()
```


# 1 Flight delay predictions

\footnotesize
\medskip
```{r}
#http://asbcllc.com/nbastatR/index.html

library(nbastatR)
library(future)
library(stringi)
library(tidyverse)
library(lubridate)
library(texreg)
library(broom)
library(knitr)
library(ggpubr)
library(ggrepel)
library(janitor)
library(plotly)

plan(multiprocess) 

# Run only when needed
# game_logs(seasons = 1947:2019, result_types = c("team", "player"))
# dataGameLogsTeam$Team = substring(dataGameLogsTeam$slugMatchup, 1, 3)

# Run when you updated data
# save(df_nba_player_dict, file='df_nba_player_dict.Rdata')
# save(dataGameLogsTeam, file='dataGameLogsTeam.Rdata')
# save(dataGameLogsPlayer, file='dataGameLogsPlayer.Rdata')

load('df_nba_player_dict.Rdata')
load('dataGameLogsTeam.Rdata')
load('dataGameLogsPlayer.Rdata')

avg <- aggregate(dataGameLogsTeam[, 24:46], list(dataGameLogsTeam$yearSeason, dataGameLogsTeam$Team), mean)
colnames(avg)[1] <- "Year"
colnames(avg)[2] <- "Team"



avgplot <- avg %>%
            filter(Team %in% c('GSW', 'CHI', 'HOU', 'LAL')) %>%
            ggplot(aes(x=Year, y=pctFG3Team, colour=Team)) +
            geom_line()
avgplot

avg2 <- aggregate(dataGameLogsTeam[, 24:46], list(dataGameLogsTeam$yearSeason), mean)
colnames(avg2)[1] <- "Year"
min2 <- aggregate(dataGameLogsTeam[, 24:46], list(dataGameLogsTeam$yearSeason), min)
colnames(min2)[1] <- "Year"
max2 <- aggregate(dataGameLogsTeam[, 24:46], list(dataGameLogsTeam$yearSeason), max)
colnames(max2)[1] <- "Year"

avgplot2 <- avg2 %>%
              filter(Year >= 1986) %>%
              ggplot(aes(x=Year, y=pctFG3Team*100)) +
              geom_path(colour='violet', size=2)
avgplot2

avgplot3 <- avg2 %>%
              ggplot(aes(x=Year, y=pctFG2Team*100)) +
              geom_path(colour='red', size=2)
avgplot3

avgminmax <- ggplot() +
                geom_line(data=min2, aes(x=Year, y=pctFG3Team*100), size=1, colour='green') +
                geom_line(data=max2, aes(x=Year, y=pctFG2Team*100), size=1, colour='red') +
                geom_smooth(data=avg2, aes(x=Year, y=pctFG3Team*100), size=2, colour='black')
avgminmax

avg1986 <- avg %>% filter(Year>=1986)
avg21986 <- avg2 %>% filter(Year>=1986)

avgcombinedall <- ggplot() +
                geom_line(data=avg1986, aes(x=Year, y=pctFG3Team*100, colour=Team), size=0.5, show.legend=FALSE) +
                geom_line(data=avg1986, aes(x=Year, y=pctFG2Team*100, colour=Team), size=0.5, show.legend=FALSE) +
                geom_line(data=avg21986, aes(x=Year, y=pctFG3Team*100), size=2, colour='black')+
                geom_line(data=avg21986, aes(x=Year, y=pctFG2Team*100), size=2, colour='black')
avgcombinedall

avgfiltered <- avg %>% filter(Team %in% c('GSW', 'CHI', 'HOU', 'LAL'))

avgcombined <- ggplot() +
                geom_line(data=avgfiltered, aes(x=Year, y=pctFG3Team*100, colour=Team), size=1, show.legend=FALSE) +
                geom_line(data=avgfiltered, aes(x=Year, y=pctFG2Team*100, colour=Team), size=1, show.legend=FALSE) +
                geom_smooth(data=avg2, aes(x=Year, y=pctFG3Team*100), size=2, colour='black')
avgcombined

avgcombined2 <- avgcombined +
                geom_smooth(data=avg2, aes(x=Year, y=pctFG2Team*100), size=2, colour='red')
avgcombined2

avgcombined3 <- avgcombined2 +
                geom_smooth(data=avg2, aes(x=Year, y=pctFTTeam*100), size=2, colour='green')
avgcombined3

avgcombined4 <- avgcombined3 +
                geom_smooth(data=avg2, aes(x=Year, y=ptsTeam), size=2, colour='purple')
avgcombined4

# ggplotly(p=ggplot2::last_plot())

# library(ggplot2)
# library(ggpubr)
# theme_set(theme_pubr())
# 
# figure <- ggarrange(avgplot, avgplot2,
#                     labels = c("Each Team", "All Teams"),
#                     ncol = 1, nrow = 2)
# figure

# climate <- read.csv('ps5_data.csv')
# a <- ggplot(climate) +
#       xlab('Year') +
#       ylab('Temperature(°C)') + 
#       theme(panel.border=element_rect(colour="black", fill=NA), panel.background=element_rect(fill=NA),
#             panel.grid=element_line(color="grey")) + 
#       geom_smooth(aes(Year, Lowess.5.), colour="blue", size=1) +
#       geom_line(aes(Year, No_Smoothing), colour="grey", size=1) + 
#       geom_point(aes(Year, No_Smoothing), shape=1, size=3)
# 

```

<!--
```{r}
# I don't know the agents
agentsPlayers <- agents_players(all_agents=TRUE, return_message = TRUE)
agents_players(agents = c("Jeff Schwartz"))

# Something like all star?
allNBATeams <- all_nba_teams()
MJTeams <- allNBATeams %>% filter(allNBATeams$namePlayer == 'Michael Jordan')

# All Star Game Information: Geo-mapping possible
library(dplyr)
library(nbastatR)
df_asg <-
all_star_games()
df_asg %>% glimpse()
df_asg %>% count(namePlayerMVP, sort = T)

# ???
assign_bref_data(dataGameLogsTeam)

assign_nba_players()
df_dict_nba_players %>% filter(namePlayer == 'Michael Jordan')

assign_nba_teams()

beyond_the_numbers(count_articles = 10)

# Box Scores - can be joined with dataGameLogsTeam
# Not working
# box_scores(game_ids = c(21800958))
# box_scores(game_ids = c(21700002, 21700003), box_score_types = c("Traditional", "Advanced", "Scoring", "Misc", "Usage", "Four Factors", "Tracking"), result_types = c("player", "team"), join_data = TRUE, assign_to_environment = TRUE, return_message = TRUE)

# Do Nothing
bref_awards(awards = c("Most Valuable Player", "Rookie of the Year", "Defensive Player of the Year"))
awards<- bref_awards(awards = c("Most Valuable Player", "Rookie of the Year",
  "Defensive Player of the Year", "Sixth Man of the Year",
  "Most Improved Player", "Teammate of the Year",
  "J. Walter Kennedy Citizenship Award", "NBA Finals Most Valuable Player",
  "ABA Playoffs Most Valuable Player",
  "All-Star Game Most Valuable Player", "Comeback Player of the Year",
  "Sporting News MVP", "Sporting News Rookie of the Year",
  "Coach of the Year", "Executive of the Year"))

# Error
bref_awards_votes()

# All players bio. can be mapped to map
bref_bios( players = c("Michael Jordan","Stephen Curry"),
player_ids = NULL,
assign_to_environment = TRUE)

brefInjuries <- bref_injuries()

brefPlayerStats <- bref_players_stats(seasons = 2017:2019, tables = c("advanced", "totals"))

brefPlayerStats %>% filter(namePlayer=='Stephen Curry')

brefPlayerStats2 <- brefPlayerStats[order(-brefPlayerStats$pct3PRate),] %>% filter(minutesTotals > 2000)

# dataBREFTeamJoined
bref_teams_stats(seasons = 2017:2018)

coachingStaffs <- coaching_staffs()

currentSchedule <- current_schedule()

# Current Standings 2018-2019 season
currentStandings <- current_standings()

# dataScore * 4
# Team Leaders
# Line Score
# Last Meeting
# Game Header
daysScores <- days_scores(game_dates = "2019-03-05", include_standings = F, return_message = T)

df_dict_awards <- dictionary_bref_awards()

df_dict_coaches <- dictionary_bref_coaches()

df_dict_players <- dictionary_bref_players()

df_dict_teams <- dictionary_bref_teams()

df_dict_column_names <- dictionary_nba_names()

df_dict_player_photos <- dictionary_player_photos()

draftCombines <- draft_combines(years = c(2001:2019), nest_data = T)

library(dplyr)
df_drafts <-
drafts(draft_years = 1983:2019, nest_data = FALSE, return_message = TRUE)

## Where do top 5 picks since 1983 come from?

df_drafts %>%
filter(numberPickOverall <= 5) %>%
count(nameOrganizationFrom, sort = T)

# Per game summary of all players
fanduelSummary <- fanduel_summary(game_ids = c(21700002), nest_data = F, return_message = T)

franchiseLeadersGSW <- franchise_leaders(teams = "Golden State Warriors", modes = c("Totals"))

# NOT RUN {
game_logs(seasons = 2019, result_types = c("team", "player"))
# }

# ???
gather_data(daysScoreTeamLeadersNBA, variable_name = "item", numeric_ids = c("^id"),
  use_logical_keys = TRUE, use_factor_keys = TRUE,
  unite_columns = NULL, separate_columns = NULL,
  use_date_keys = FALSE, remove_na = TRUE)

hoops_hype_salary <- hoops_hype_salary_summary()

# Error
library(dplyr)
df_salaries <-
hoopshype_salaries(all_teams = TRUE, nest_data = F, return_message = T)
 ## By Expiring Salary Type and Team
 df_salaries %>% group_by(slugSeason,  nameTeam, isFinalSeason) %>%
 summarise(expiringSalaries = sum(amountContractMil, na.rm = T) / 1000000) %>%
 arrange(nameTeam)

# Relationship between points and 3pt
metricsLeader <- metrics_leaders(seasons = 2000:2019,
 metric = "pts",
 season_types = "Regular Season",
 modes = "Totals",
 return_message = T)

franchiseHistory <- nba_franchise_history()

# salaries & 3pt
nbaInsiderSalaries <- nba_insider_salaries(assume_player_opt_out = T, assume_team_doesnt_exercise = T, return_message = TRUE)

nba_player_ids(players = c("Mitch Richmond", "Kyle Kuzma"))

# nbaPlayers == df_dict_nba_players
nbaPlayers <- nba_players()

# Error
nba_stats_api_items()

# nbaTeams == df_dict_nba_teams
nbaTeams <- nba_teams()

nba_teams_ids(teams = c("Brooklyn Nets", "Denver Nuggets"))

nbaTeamsSeasons <- nba_teams_seasons()

nbadraftnet <- nbadraftnet_mock_drafts(years = 2012:2017,
  merge_nba_data = TRUE,
  nest_data = F,
  return_message = T)

nbastatsAPIParameters <- nbastats_api_parameters(1:3)

playByPlay <- play_by_play(game_ids = c(21700002), nest_data = F, return_message = T)

playerProfiles <- player_profiles(player_ids = c(203500, 1628384),
  players = c("Michael Jordan", "Caris LeVert", "Jarrett Allen", "James Harden"),
  nest_data = FALSE,
  return_message = TRUE)

library(dplyr)
df_players_agents <- players_agents()
df_players_agents %>% glimpse()

playersAwards <- players_awards(players = c( "Charles Oakley", "Gary Melchionni"),
  player_ids = c(893, 76375),
  return_message = T,
  nest_data = F)

playersBios <- players_bios(players = c("Michael Jordan", "Caris LeVert", "Jarrett Allen", "James Harden"))

playersCareers <- players_careers(players = c("Joe Harris", "Myles Turner", "Spencer Dinwiddie"),
modes = c("Totals", "PerGame"))

players_rotowire(players = c( "Jarrett Allen", "DeMarre Carroll", "Allen Crabbe"))

playersTables <- players_tables(players = c("Caris LeVert", "Joe Harris"), tables =  c("year over year", "passes", "game splits"),   modes = c("PerGame", "Totals"), measures = c("Base", "Advanced"), assign_to_environment = TRUE)

playoffPictures <- playoff_pictures(seasons = 2015:2018,
  assign_to_environment = TRUE,
  include_numeric_records = T)

# Error
pst_transaction(person = NULL, team = "Nets")

# Summarize data per minute?
scale_per_minute(data, scale_columns = NULL)

seasonPlayers <- seasons_players(2010:2017, nest_data = T, return_message = T)

# Error
df_rosters <- seasons_rosters(2015:2018)

### Mean Age by Season and Team
df_rosters %>%
  group_by(slugSeason, slugTeam) %>%
  summarise(ageMean = mean(agePlayer)) %>%
  arrange(ageMean) %>%
  ungroup()

# Error
seasonsSchedule <- seasons_schedule(seasons = c(2012, 2018))

# Summer League
sl_players()
sl_teams()

spread_data(data, variable_name = "item", value_name = "value",
  perserve_order = TRUE, unite_columns = NULL,
  separate_columns = NULL)

# Error
standings <- standings(seasons = 2015:2018, season_types = "Regular Season", resolve_records = T, nest_data = F, return_message = T)

summarise_per_minute(data, id_columns = c("idPlayerSeason"),
  scale_columns = c("pts", "fg", "ast", "tov", "blk", "stl", "drb",
  "trb", "orb", "ft", "pf", "countLayupsShooting", "countDunks", "hlf"))

# Error
synergy(seasons = 2019, result_types = c("player", "team"), season_types = c("Regular Season"), set_types = c("offensive", "defensive"), categories = c("Transition", "Isolation", "PRBallHandler", "PRRollman", "Postup",  "Spotup", "Handoff", "Cut", "OffScreen", "OffRebound", "Misc"), results = 500, assign_to_environment = TRUE, return_wide = F, return_message = TRUE)

# Error
team_season_roster(team = "Denver Nuggets", season = 1991)

teamsAnnualStats <- teams_annual_stats(all_active_teams = T,
  modes = c("Totals"),
  return_message = TRUE,
  nest_data =F)

# Error
teamsCoaches <- teams_coaches(2018:2019)

# Error
teamsDetails <- teams_details(all_teams = TRUE, assign_to_environment = TRUE)

teamsPlayersStats <- teams_players_stats(seasons = 2018, types = c("player", "team"),
 modes = c("PerGame", "Totals"),
 tables = c("general", "defense", "clutch", "hustle", "shots", "shot locations"))

# interesting team rankings
teamsRankings <- teams_rankings(seasons = 2019)

# too long
teamsRosters <- teams_rosters(seasons = 2010:2018, nest_data = F, return_message = T)

# Error
teamsRotowire <- teams_rotowire(teams = "Brooklyn Nets")

# too long
teamsSeasonsInfo <- teams_seasons_info(teams = "Brooklyn Nets", seasons = c(1984, 1990, 1995, 2018), season_types = "Regular Season")

# too long
teamsShots <- teams_shots(teams = "Golden State Warriors", seasons = 2018)
teamsTables <- teams_tables(teams = c("Brooklyn Nets", "New York Knicks"),
 seasons = 2017:2018, tables = c("splits", "shooting"), measures = "Base", modes = c("PerGame", "Totals"))

transactions <- transactions()

validate_nba_player_photos(sleep_time = 2)

widen_bref_data(data)

win_probability(game_ids = c(21700002, 21700005), filter_non_plays = T,
nest_data = FALSE,
return_message = TRUE)





```
-->
\normalsize


-	It seems that players are getting better at making 3-pointers than 10 or 20 years ago (both on average and also top 3-pointer shooters vs. top 3-pointer shooters) Is it true?

```{r}
avgplot2
```

-	If true, what could be the reasons for that?

-	Teams with more 3-pointers tend to be the better performing teams?

-	Are there any relationship between players’ ages and 3-pointers? Both total and average.

-	Players who are good at 3-pointers are also good at 2-pointers or free throws?

-	Any insights between standings and 3-pointers?

-	Players with high salaries are good at 3-pointers?

-	We want to analyze whether players can drastically improve their three point shooting skills over time or the skill is rather something people are borned with.

-	Show the 3-pointer statistics geographically based on players’ hometowns. Maybe this help illustrates the different basketball playing style across different regions, both domestic and international.

-	We would like to explore the importance of three point shooters in a given team by measuring the share of the team’s total salary over time.

-	Focus on three point shooting is a strategy that started fairly recently, we can create a map to show where this strategy initially emerged and how fast it spreaded across the entire country.

-	What are the expected average points of 3-pointers and 2-pointers? Show the historical data.

-	If the expected average point from 3-pointers is getting higher than that of 2-pointers, how should each team’s strategy changes

