---
title: "Problem Set 6"
author: "Subeom Lee"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    df_print: paged
linkcolor: cyan
fontsize: 10pt
colorlinks: yes
---

```{r, message=FALSE, echo=FALSE}
library(lubridate)
library(texreg)
library(broom)
library(knitr)
library(ggpubr)
library(ggrepel)
library(janitor)
library(reticulate)
opts_chunk$set(echo=TRUE,
               cache.lazy=FALSE,
               collapse=TRUE,
               comment=NA,
               warning=FALSE,
               prompt=FALSE,
               message=FALSE
               ,cache=TRUE
               )
options(digits=4, scipen=0)
masterdoc <- knitr::current_input()
```


# Questions

\footnotesize
\medskip
```{r}
#http://asbcllc.com/nbastatR/index.html

library(nbastatR)
library(future)
library(stringi)
library(tidyverse)
library(lubridate)
library(texreg)
library(broom)
library(knitr)
library(ggpubr)
library(ggrepel)
library(janitor)
library(plotly)
library(reticulate)

plan(multiprocess) 

# Run only when needed
# game_logs(seasons = 1947:2019, result_types = c("team", "player"))
# dataGameLogsTeam$Team = substring(dataGameLogsTeam$slugMatchup, 1, 3)

# Run when you updated data
# save(df_nba_player_dict, file='df_nba_player_dict.Rdata')
# save(dataGameLogsTeam, file='dataGameLogsTeam.Rdata')
# save(dataGameLogsPlayer, file='dataGameLogsPlayer.Rdata')

# load('df_nba_player_dict.Rdata')
# load('dataGameLogsTeam.Rdata')
# load('dataGameLogsPlayer.Rdata')

load('BaseEnvironment.Rdata')



# avg <- aggregate(dataGameLogsTeam[, 24:46], list(dataGameLogsTeam$yearSeason, dataGameLogsTeam$Team), mean)
# colnames(avg)[1] <- "Year"
# colnames(avg)[2] <- "Team"
# 
# 
# 
# avgplot <- avg %>%
#             filter(Team %in% c('GSW', 'CHI', 'HOU', 'LAL')) %>%
#             ggplot(aes(x=Year, y=pctFG3Team, colour=Team)) +
#             geom_line()
# avgplot
# 
# avg2 <- aggregate(dataGameLogsTeam[, 24:46], list(dataGameLogsTeam$yearSeason), mean)
# colnames(avg2)[1] <- "Year"
# min2 <- aggregate(dataGameLogsTeam[, 24:46], list(dataGameLogsTeam$yearSeason), min)
# colnames(min2)[1] <- "Year"
# max2 <- aggregate(dataGameLogsTeam[, 24:46], list(dataGameLogsTeam$yearSeason), max)
# colnames(max2)[1] <- "Year"
# 
# avgplot2 <- avg2 %>%
#               filter(Year >= 1986) %>%
#               ggplot(aes(x=Year, y=pctFG3Team*100)) +
#               geom_path(colour='violet', size=2)
# avgplot2
# 
# avgplot3 <- avg2 %>%
#               ggplot(aes(x=Year, y=pctFG2Team*100)) +
#               geom_path(colour='red', size=2)
# avgplot3
# 
# avgminmax <- ggplot() +
#                 geom_line(data=min2, aes(x=Year, y=pctFG3Team*100), size=1, colour='green') +
#                 geom_line(data=max2, aes(x=Year, y=pctFG2Team*100), size=1, colour='red') +
#                 geom_smooth(data=avg2, aes(x=Year, y=pctFG3Team*100), size=2, colour='black')
# avgminmax
# 
# avg1986 <- avg %>% filter(Year>=1986)
# avg21986 <- avg2 %>% filter(Year>=1986)
# 
# avgcombinedall <- ggplot() +
#                 geom_line(data=avg1986, aes(x=Year, y=pctFG3Team*100, colour=Team), size=0.5, show.legend=FALSE) +
#                 geom_line(data=avg1986, aes(x=Year, y=pctFG2Team*100, colour=Team), size=0.5, show.legend=FALSE) +
#                 geom_line(data=avg21986, aes(x=Year, y=pctFG3Team*100), size=2, colour='black')+
#                 geom_line(data=avg21986, aes(x=Year, y=pctFG2Team*100), size=2, colour='black')
# avgcombinedall
# 
# avgfiltered <- avg %>% filter(Team %in% c('GSW', 'CHI', 'HOU', 'LAL'))
# 
# avgcombined <- ggplot() +
#                 geom_line(data=avgfiltered, aes(x=Year, y=pctFG3Team*100, colour=Team), size=1, show.legend=FALSE) +
#                 geom_line(data=avgfiltered, aes(x=Year, y=pctFG2Team*100, colour=Team), size=1, show.legend=FALSE) +
#                 geom_smooth(data=avg2, aes(x=Year, y=pctFG3Team*100), size=2, colour='black')
# avgcombined
# 
# avgcombined2 <- avgcombined +
#                 geom_smooth(data=avg2, aes(x=Year, y=pctFG2Team*100), size=2, colour='red')
# avgcombined2
# 
# avgcombined3 <- avgcombined2 +
#                 geom_smooth(data=avg2, aes(x=Year, y=pctFTTeam*100), size=2, colour='green')
# avgcombined3
# 
# avgcombined4 <- avgcombined3 +
#                 geom_smooth(data=avg2, aes(x=Year, y=ptsTeam), size=2, colour='purple')
# avgcombined4

# ggplotly(p=ggplot2::last_plot())

# library(ggplot2)
# library(ggpubr)
# theme_set(theme_pubr())
# 
# figure <- ggarrange(avgplot, avgplot2,
#                     labels = c("Each Team", "All Teams"),
#                     ncol = 1, nrow = 2)
# figure

# climate <- read.csv('ps5_data.csv')
# a <- ggplot(climate) +
#       xlab('Year') +
#       ylab('Temperature(°C)') + 
#       theme(panel.border=element_rect(colour="black", fill=NA), panel.background=element_rect(fill=NA),
#             panel.grid=element_line(color="grey")) + 
#       geom_smooth(aes(Year, Lowess.5.), colour="blue", size=1) +
#       geom_line(aes(Year, No_Smoothing), colour="grey", size=1) + 
#       geom_point(aes(Year, No_Smoothing), shape=1, size=3)
# 

```
\normalsize

# Team level questions

Q1.	It seems that players are getting better at making 3-pointers than 20 years ago (both on average and also top 3-pointer shooters vs. top 3-pointer shooters) Is it true?

```{r}
fg3year <- aggregate(dataGameLogsTeam[, 35:36], list(dataGameLogsTeam$yearSeason), sum)
colnames(fg3year)[1] <- "Year"
fg3year <- fg3year %>% filter (Year >= 1986)
fg3year$pctfg3 <- fg3year$fg3mTeam / fg3year$fg3aTeam * 100

fg3yearteam <- aggregate(dataGameLogsTeam[, 35:36], list(dataGameLogsTeam$yearSeason, dataGameLogsTeam$Team), sum)
colnames(fg3yearteam)[1] <- "Year"
colnames(fg3yearteam)[2] <- "Team"
fg3yearteam <- fg3yearteam %>% filter (Year >= 1986)
fg3yearteam$pctfg3 <- fg3yearteam$fg3mTeam / fg3yearteam$fg3aTeam * 100

xaxisbreaks <- seq(1985, 2019, by=3)
yaxisbreaks <- seq(15, 45, by=5)

Q1 <- ggplot() +
  geom_line(data=fg3yearteam, aes(x=Year, y=pctfg3, colour=Team), size=0.5, show.legend=FALSE, alpha=0.7) +
  geom_line(data=fg3year, aes(x=Year, y=pctfg3), size=2, colour='black') +
  xlab('Year') +
  ylab(NULL) +
  ggtitle('3 Pointer Field Goal Success Rate') +
  theme(panel.background=element_rect(fill=NA), panel.grid.major.y=element_line(color="grey", linetype=2),
        plot.title = element_text(hjust = 0.5)) + 
  scale_y_continuous(limits=c(15, 45), breaks=yaxisbreaks, labels=paste(yaxisbreaks,"%")) +
  scale_x_continuous(limits=c(1985,2019), breaks=xaxisbreaks) +
  geom_hline(yintercept=min(fg3year$pctfg3), linetype=2, color="steelblue", size=0.5, alpha=0.9) +
  geom_hline(yintercept=max(fg3year$pctfg3), linetype=2, color="steelblue", size=0.5, alpha=0.9) +
  geom_hline(yintercept=min(fg3yearteam$pctfg3), linetype=3, color="pink", size=0.5, alpha=0.9) +
  geom_hline(yintercept=max(fg3yearteam$pctfg3), linetype=3, color="pink", size=0.5, alpha=0.9) +
  annotate("text", x=1985, y=min(fg3year$pctfg3)+0.6, label=paste(toString(round(min(fg3year$pctfg3), digits=2)),"%"), color="steelblue", size=3) +
  annotate("text", x=1985, y=max(fg3year$pctfg3)+0.6, label=paste(toString(round(max(fg3year$pctfg3), digits=2)),"%"), color="steelblue", size=3) +
  annotate("text", x=1985, y=min(fg3yearteam$pctfg3)+0.6, label=paste(toString(round(min(fg3yearteam$pctfg3), digits=2)),"%"), color="pink", size=3) +
  annotate("text", x=1985, y=max(fg3yearteam$pctfg3)+0.6, label=paste(toString(round(max(fg3yearteam$pctfg3), digits=2)),"%"), color="pink", size=3)  
  
Q1


fg3yearavg <- aggregate(dataGameLogsTeam[, 35:36], list(dataGameLogsTeam$yearSeason), mean)
colnames(fg3yearavg)[1] <- "Year"
fg3yearavg <- fg3yearavg %>% filter (Year >= 1986)
fg3yearavg$pctfg3 <- fg3yearavg$fg3mTeam / fg3yearavg$fg3aTeam * 100

fg3yearteamavg <- aggregate(dataGameLogsTeam[, 35:36], list(dataGameLogsTeam$yearSeason, dataGameLogsTeam$Team), mean)
colnames(fg3yearteamavg)[1] <- "Year"
colnames(fg3yearteamavg)[2] <- "Team"
fg3yearteamavg <- fg3yearteamavg %>% filter (Year >= 1986)
fg3yearteamavg$pctfg3 <- fg3yearteamavg$fg3mTeam / fg3yearteamavg$fg3aTeam * 100

xaxisbreaks <- seq(1986, 2019, by=3)
yaxisbreaks <- seq(0, 50, by=3)

Q1_2 <- ggplot() +
  geom_line(data=fg3yearteamavg, aes(x=Year, y=fg3mTeam, colour=Team), size=0.5, show.legend=FALSE, alpha=0.7) +
  geom_line(data=fg3yearavg, aes(x=Year, y=fg3mTeam), size=2, colour='green') +
  geom_line(data=fg3yearteamavg, aes(x=Year, y=fg3aTeam, colour=Team), size=0.5, show.legend=FALSE, alpha=0.7) +
  geom_line(data=fg3yearavg, aes(x=Year, y=fg3aTeam), size=2, colour='blue') +
  geom_line(data=fg3year, aes(x=Year, y=pctfg3), size=2, colour='black') +
  xlab('Year') +
  ylab(NULL) +
  ggtitle('3 Pointer Field Goal made vs tries') +
  theme(panel.background=element_rect(fill=NA), panel.grid.major.y=element_line(color="grey", linetype=2),
        plot.title = element_text(hjust = 0.5)) + 
  scale_y_continuous(limits=c(0, 50), breaks=yaxisbreaks) +
  scale_x_continuous(limits=c(1986,2019), breaks=xaxisbreaks)
  
Q1_2


fgallyearavg <- aggregate(dataGameLogsTeam[, 29:38], list(dataGameLogsTeam$yearSeason), mean)
colnames(fgallyearavg)[1] <- "Year"
fgallyearavg["plusminusTeam"] = NULL
fgallyearavg["urlTeamSeasonLogo"] = NULL
fgallyearavg["pfTeam"] = NULL
fgallyearavg <- fgallyearavg %>% filter (Year >= 1986)
fgallyearavg$pctpts3 <- fgallyearavg$fg3mTeam / fgallyearavg$ptsTeam * 100
fgallyearavg$pctpts2 <- fgallyearavg$fg2mTeam / fgallyearavg$ptsTeam * 100
fgallyearavg$pctptsft <- fgallyearavg$ftmTeam / fgallyearavg$ptsTeam * 100

fgallyearteamavg <- aggregate(dataGameLogsTeam[, 29:38], list(dataGameLogsTeam$yearSeason, dataGameLogsTeam$Team), mean)
colnames(fgallyearteamavg)[1] <- "Year"
colnames(fgallyearteamavg)[2] <- "Team"
fgallyearteamavg["plusminusTeam"] = NULL
fgallyearteamavg["urlTeamSeasonLogo"] = NULL
fgallyearteamavg["pfTeam"] = NULL
fgallyearteamavg <- fgallyearteamavg %>% filter (Year >= 1986)
fgallyearteamavg$pctpts3 <- fgallyearteamavg$fg3mTeam / fgallyearteamavg$ptsTeam * 100
fgallyearteamavg$pctpts2 <- fgallyearteamavg$fg2mTeam / fgallyearteamavg$ptsTeam * 100
fgallyearteamavg$pctptsft <- fgallyearteamavg$ftmTeam / fgallyearteamavg$ptsTeam * 100

xaxisbreaks <- seq(1986, 2019, by=3)
yaxisbreaks <- seq(0, 45, by=5)

Q1_3 <- ggplot() +
  geom_line(data=fgallyearteamavg, aes(x=Year, y=pctpts3, colour=Team), size=0.5, show.legend=FALSE, alpha=0.7) +
  geom_line(data=fgallyearteamavg, aes(x=Year, y=pctpts2, colour=Team), size=0.5, show.legend=FALSE, alpha=0.7) +
  geom_line(data=fgallyearteamavg, aes(x=Year, y=pctptsft, colour=Team), size=0.5, show.legend=FALSE, alpha=0.7) +
  geom_line(data=fgallyearavg, aes(x=Year, y=pctpts3), size=2, colour='red') +
  geom_line(data=fgallyearavg, aes(x=Year, y=pctpts2), size=2, colour='green') +
  geom_line(data=fgallyearavg, aes(x=Year, y=pctptsft), size=2, colour='blue') +
  xlab('Year') +
  ylab(NULL) +
  ggtitle('Field Goal Percentage / all Points red:3, green: 2, blue: free throws') +
  theme(panel.background=element_rect(fill=NA), panel.grid.major.y=element_line(color="grey", linetype=2),
        plot.title = element_text(hjust = 0.5)) + 
  scale_y_continuous(limits=c(0, 45), breaks=yaxisbreaks) +
  scale_x_continuous(limits=c(1986,2019), breaks=xaxisbreaks)
  
Q1_3
```

Yes, the success rate of 3 point field goal has been increased by about 9% since 1986.

Q2.	If true, what could be the reasons for that?
    -	What are the expected average points of 3-pointers and 2-pointers? Show the historical data.
    -	If the expected average point from 3-pointers is getting higher than that of
      2-pointers, how should each team’s strategy changes

https://www.nytimes.com/2016/01/21/sports/basketball/how-the-nba-3-point-shot-went-from-gimmick-to-game-changer.html

Its debut, in the 1979-80 season, was inauspicious.

There are many reasons for the rise of the 3-point shot, but one may simply be math. It took a while, but coaches finally stopped listening to the traditionalist naysayers and realized that a shot that is worth 50 percent more pays off, even if that shot is a little harder to make.

“Teams have all caught on to the whole points-per-possession argument,” Lawrence Frank, the Nets’ coach at the time, said in 2009 as the 3 rate began to rapidly increase.

```{r}
fgyear <- aggregate(dataGameLogsTeam[, 35:38], list(dataGameLogsTeam$yearSeason), sum)
colnames(fgyear)[1] <- "Year"
fgyear <- fgyear %>% filter (Year >= 1986)
fgyear$pctfg3 <- fgyear$fg3mTeam / fgyear$fg3aTeam * 100
fgyear$pctfg2 <- fgyear$fg2mTeam / fgyear$fg2aTeam * 100

fgyearteam <- aggregate(dataGameLogsTeam[, 35:38], list(dataGameLogsTeam$yearSeason, dataGameLogsTeam$Team), sum)
colnames(fgyearteam)[1] <- "Year"
colnames(fgyearteam)[2] <- "Team"
fgyearteam <- fgyearteam %>% filter (Year >= 1986)
fgyearteam$pctfg3 <- fgyearteam$fg3mTeam / fgyearteam$fg3aTeam * 100
fgyearteam$pctfg2 <- fgyearteam$fg2mTeam / fgyearteam$fg2aTeam * 100

xaxisbreaks <- seq(1985, 2019, by=3)
yaxisbreaks <- seq(15, 60, by=5)

Q2_1 <- ggplot() +
  geom_line(data=fgyearteam, aes(x=Year, y=pctfg3, colour=Team), size=0.5, show.legend=FALSE, alpha=0.7) +
  geom_line(data=fgyear, aes(x=Year, y=pctfg3), size=2, colour='black') +
  geom_line(data=fgyearteam, aes(x=Year, y=pctfg2, colour=Team), size=0.5, show.legend=FALSE, alpha=0.7) +
  geom_line(data=fgyear, aes(x=Year, y=pctfg2), size=2, colour='black') +
  xlab('Year') +
  ylab(NULL) +
  ggtitle('Field Goal Success Rate') +
  theme(panel.background=element_rect(fill=NA), panel.grid.major.y=element_line(color="grey", linetype=2),
        plot.title = element_text(hjust = 0.5)) + 
  scale_y_continuous(limits=c(15, 60), breaks=yaxisbreaks, labels=yaxisbreaks) +
  scale_x_continuous(limits=c(1985,2019), breaks=xaxisbreaks)# +
  # geom_hline(yintercept=min(fg3year$pctfg3), linetype=2, color="steelblue", size=0.5, alpha=0.9) +
  # geom_hline(yintercept=max(fg3year$pctfg3), linetype=2, color="steelblue", size=0.5, alpha=0.9) +
  # geom_hline(yintercept=min(fg3yearteam$pctfg3), linetype=3, color="pink", size=0.5, alpha=0.9) +
  # geom_hline(yintercept=max(fg3yearteam$pctfg3), linetype=3, color="pink", size=0.5, alpha=0.9) +
  # annotate("text", x=1985, y=min(fg3year$pctfg3)+0.6, label=paste(toString(round(min(fg3year$pctfg3), digits=2)),"%"), color="steelblue", size=3) +
  # annotate("text", x=1985, y=max(fg3year$pctfg3)+0.6, label=paste(toString(round(max(fg3year$pctfg3), digits=2)),"%"), color="steelblue", size=3) +
  # annotate("text", x=1985, y=min(fg3yearteam$pctfg3)+0.6, label=paste(toString(round(min(fg3yearteam$pctfg3), digits=2)),"%"), color="pink", size=3) +
  # annotate("text", x=1985, y=max(fg3yearteam$pctfg3)+0.6, label=paste(toString(round(max(fg3yearteam$pctfg3), digits=2)),"%"), color="pink", size=3)  
  # 
Q2_1

```

The expected points of 2-point shots in 1986 was 'r fgyear$pctfg2[1986-1985]/100' * 2 = 'r fgyear$pctfg2[1986-1985]/100*2'
The expected points of 3-point shots in 1986 was 'r fgyear$pctfg3[1986-1985]/100' * 3 = 'r fgyear$pctfg3[1986-1985]/100*3'

The expected points of 2-point shots in 2019 was 'r fgyear$pctfg2[2019-1985]/100' * 2 = 'r fgyear$pctfg2[2019-1985]/100*2'
The expected points of 3-point shots in 2019 was 'r fgyear$pctfg3[2019-1985]/100' * 3 = 'r fgyear$pctfg3[2019-1985]/100*3'

Teams started to focus on 3-point shots after its first introduction in 1979,
because the expected points of 3-point shots are higher than that of 2-point shots
since early 90's.

```{r}
fgyear$e2 = fgyear$pctfg2 / 100 * 2
fgyear$e3 = fgyear$pctfg3 / 100 * 3

xaxisbreaks <- seq(1985, 2019, by=3)
yaxisbreaks <- seq(0.8, 1.15, by=0.05)

Q2_2 <- ggplot() +
  geom_line(data=fgyear, aes(x=Year, y=e3), size=1, colour='red') +
  geom_line(data=fgyear, aes(x=Year, y=e2), size=1, colour='blue') +
  xlab('Year') +
  ylab(NULL) +
  ggtitle('Expected points') +
  theme(panel.background=element_rect(fill=NA), panel.grid.major.y=element_line(color="grey", linetype=2),
        plot.title = element_text(hjust = 0.5)) + 
  scale_y_continuous(limits=c(0.8, 1.15), breaks=yaxisbreaks, labels=yaxisbreaks) +
  scale_x_continuous(limits=c(1985,2019), breaks=xaxisbreaks)

Q2_2

```

Q3.	Teams with more 3-pointers tend to be the better performing teams?
    -	Any insights between standings and 3-pointers?


```{r}
standings <- read_csv("standings.csv")

fgyearteam <- aggregate(dataGameLogsTeam[, 35:38], list(dataGameLogsTeam$yearSeason, dataGameLogsTeam$nameTeam), sum)
colnames(fgyearteam)[1] <- "Year"
colnames(fgyearteam)[2] <- "nameTeam"
fgyearteam <- fgyearteam %>% filter (Year >= 1986)
fgyearteam$pctfg3 <- fgyearteam$fg3mTeam / fgyearteam$fg3aTeam * 100
fgyearteam$pctfg2 <- fgyearteam$fg2mTeam / fgyearteam$fg2aTeam * 100

standings2 <- left_join(standings, fgyearteam, by=c("Year" = "Year", "Team" = "nameTeam"))

Q3 <- ggplot(standings2) +
  geom_point(aes(x=pctfg3, y=Rk), color="steelblue") +
  geom_point(aes(x=pctfg2, y=Rk), color="pink")
  # geom_line(data=fgyear, aes(x=Year, y=e2), size=1, colour='blue') +
  # xlab('Year') +
  # ylab(NULL) +
  # ggtitle('Expected points') +
  # theme(panel.background=element_rect(fill=NA), panel.grid.major.y=element_line(color="grey", linetype=2),
  #       plot.title = element_text(hjust = 0.5)) + 
  # scale_y_continuous(limits=c(0.8, 1.15), breaks=yaxisbreaks, labels=yaxisbreaks) +
  # scale_x_continuous(limits=c(1985,2019), breaks=xaxisbreaks)
  
Q3

linearModel <- lm(Rk ~ pctfg3, data=standings2)
tidy(linearModel)

linearModel2 <- lm(Rk ~ pctfg2, data=standings2) 
tidy(linearModel2)

linearModel3 <- lm(Rk ~ pctfg3 + pctfg2, data=standings2)
tidy(linearModel3)

# plot(linearModel)
# plot(linearModel2)
# plot(linearModel3)

linearModel4 <- lm(pctfg3 ~ pctfg2, data=standings2)
tidy(linearModel4)
# plot(linearModel4)

Q3_2 <- ggplot(standings2) +
  geom_point(aes(x=pctfg2, y=pctfg3), color="steelblue")
Q3_2
 
```

Yes. However, pctfg2 is more relevant than pctfg3



-	Focus on three point shooting is a strategy that started fairly recently, we can create a map to show where this strategy initially emerged and how fast it spreaded across the entire country.



# Player level questions

```{r}
dataGameLogsPlayer1986 <- dataGameLogsPlayer %>% filter(yearSeason >= 1986)

fgyearplayer <- aggregate(dataGameLogsPlayer1986[, 19:26], list(dataGameLogsPlayer1986$yearSeason, dataGameLogsPlayer1986$namePlayer), sum)
colnames(fgyearplayer)[1] <- "Year"
colnames(fgyearplayer)[2] <- "Player"
fgyearplayer$pctFG = NULL
fgyearplayer$pctFG3 = NULL

fgyearplayer$pctfg3 <- fgyearplayer$fg3m / fgyearplayer$fg3a * 100
fgyearplayer$pctfg2 <- fgyearplayer$fgm / fgyearplayer$fga * 100
fgyearplayer$pctft <- fgyearplayer$ftm / fgyearplayer$fta * 100

# Meaningless...
yearplayer <- aggregate(fgyearplayer[,5], list(fgyearplayer$Player), sum)
colnames(yearplayer)[1] <- "Player"
colnames(yearplayer)[2] <- "totalfg3m"
ggplot(yearplayer, aes(totalfg3m)) + geom_histogram()
yearplayer100 <- yearplayer %>% filter (totalfg3m>=100)

xaxisbreaks <- seq(1985, 2019, by=3)
yaxisbreaks <- seq(20, 50, by=5)

fgyearplayer100 <- fgyearplayer %>% filter(Player %in% yearplayer100$Player)
plotYearPlayer <- ggplot() +
  geom_line(data=fgyearplayer100, aes(x=Year, y=pctfg3, colour=Player), size=1, show.legend = FALSE) +
  geom_line(data=fgyearplayer100, aes(x=Year, y=pctfg2, colour=Player), size=1, show.legend = FALSE) +
  xlab('Year') +
  ylab(NULL) +
  ggtitle('3 point shot success rate by player') +
  theme(panel.background=element_rect(fill=NA), panel.grid.major.y=element_line(color="grey", linetype=2),
        plot.title = element_text(hjust = 0.5)) + 
  scale_y_continuous(limits=c(0, 100), breaks=yaxisbreaks, labels=yaxisbreaks) +
  scale_x_continuous(limits=c(1985,2019), breaks=xaxisbreaks)
plotYearPlayer
# Meaningless...

fgplayer <- aggregate(dataGameLogsPlayer1986[, 19:26], list(dataGameLogsPlayer1986$namePlayer), sum)
colnames(fgplayer)[1] <- "Player"
fgplayer$pctFG = NULL
fgplayer$pctFG3 = NULL

fgplayer$pctfg3 <- fgplayer$fg3m / fgplayer$fg3a * 100
fgplayer$pctfg2 <- fgplayer$fgm / fgplayer$fga * 100
fgplayer$pctft <- fgplayer$ftm / fgplayer$fta * 100

fgplayer <- fgplayer[order(-fgplayer$pctfg3),]
fgplayer100 <- fgplayer %>% filter(fg3m >= 100)

```

```{python}
import pandas as pd

fgplayer = r.fgplayer

fgplayer['firstYear'] = 2019
fgplayer['lastYear'] = 1986

print(fgplayer.head(5))
print(fgplayer.tail(5))

i=0

for player in fgplayer.values:
  min = player[-2]
  max = player[-1]
  for yp in r.fgyearplayer.values:
    if player[0] == yp[1]:
      if max < yp[0]: max = yp[0]
      if min > yp[0]: min = yp[0]
  fgplayer.iloc[i,-1]=max
  fgplayer.iloc[i,-2]=min
  i += 1
  
print(fgplayer.head(5))
print(fgplayer.tail(5))

```

```{r}
fgplayer <- py$fgplayer
fgplayer100 <- fgplayer %>% filter(fg3m >= 100)
fgplayer1000 <- fgplayer100 %>% filter(fg3m >= 1000)
fgplayer2000 <- fgplayer1000 %>% filter(fg3m >= 2000)


xaxisbreaks <- seq(1985, 2019, by=3)
yaxisbreaks <- seq(20, 50, by=5)

plotPlayer100 <- ggplot() +
  geom_linerange(data=fgplayer100, aes(x=pctfg3, y=lastYear, ymin=firstYear, ymax=lastYear, colour=Player), size=1, show.legend = FALSE) +
  # geom_point(data=fgplayer100, aes(x=lastYear, y=pctfg3, colour=Player), size=1, show.legend = FALSE) +
  # geom_line
  xlab('Year') +
  ylab(NULL) +
  ggtitle('3 point success rate by player and year') +
  theme(panel.background=element_rect(fill=NA), panel.grid.major.y=element_line(color="grey", linetype=2),
        plot.title = element_text(hjust = 0.5)) + 
  scale_x_continuous(limits=c(20, 50), breaks=yaxisbreaks, labels=yaxisbreaks) +
  scale_y_continuous(limits=c(1985,2019), breaks=xaxisbreaks)
plotPlayer100


xaxisbreaks <- seq(1985, 2019, by=3)
yaxisbreaks <- seq(30, 45, by=1)

fgplayer1000 <- fgplayer100 %>% filter (fg3m >= 1000)
plotPlayer1000 <- ggplot() +
  geom_pointrange(data=fgplayer1000, aes(x=pctfg3, y=lastYear, ymin=firstYear, ymax=lastYear, colour=Player), size=1, show.legend = FALSE) +
  # geom_point(data=fgplayer100, aes(x=lastYear, y=pctfg3, colour=Player), size=1, show.legend = FALSE) +
  # geom_line
  xlab('Year') +
  ylab(NULL) +
  ggtitle('3 point success rate by player and year') +
  theme(panel.background=element_rect(fill=NA), panel.grid.major.y=element_line(color="grey", linetype=2),
        plot.title = element_text(hjust = 0.5)) + 
  scale_x_continuous(limits=c(30, 45), breaks=yaxisbreaks, labels=yaxisbreaks) +
  scale_y_continuous(limits=c(1985,2019), breaks=xaxisbreaks)
plotPlayer1000


```

```{r}
xaxisbreaks <- seq(1985, 2019, by=3)
yaxisbreaks <- seq(30, 45, by=1)

plotPlayer100 <- ggplot() +
  geom_point(data=fgplayer100, aes(x=lastYear, y=pctfg3, size=fg3m+fg3a, colour=Player), show.legend = FALSE) +
  # geom_point(data=fgplayer100, aes(x=lastYear, y=pctfg3, colour=Player), size=1, show.legend = FALSE) +
  # geom_line
  xlab('Year') +
  ylab(NULL) +
  ggtitle('3 point success rate by player and year') +
  theme(panel.background=element_rect(fill=NA), panel.grid.major.y=element_line(color="grey", linetype=2),
        plot.title = element_text(hjust = 0.5)) + 
  scale_y_continuous(limits=c(30, 45), breaks=yaxisbreaks, labels=yaxisbreaks) +
  scale_x_continuous(limits=c(1985,2019), breaks=xaxisbreaks)
plotPlayer100

plotPlayer1000 <- ggplot() +
  geom_point(data=fgplayer1000, aes(x=lastYear, y=pctfg3, size=fg3m+fg3a, colour=Player), show.legend = FALSE) +
  # geom_point(data=fgplayer100, aes(x=lastYear, y=pctfg3, colour=Player), size=1, show.legend = FALSE) +
  # geom_line
  xlab('Year') +
  ylab(NULL) +
  ggtitle('3 point success rate by player and year') +
  theme(panel.background=element_rect(fill=NA), panel.grid.major.y=element_line(color="grey", linetype=2),
        plot.title = element_text(hjust = 0.5)) + 
  scale_y_continuous(limits=c(30, 45), breaks=yaxisbreaks, labels=yaxisbreaks) +
  scale_x_continuous(limits=c(1985,2019), breaks=xaxisbreaks)
plotPlayer1000

plotPlayer2000 <- ggplot() +
  geom_point(data=fgplayer2000, aes(x=lastYear, y=pctfg3, size=fg3m+fg3a, colour=Player), show.legend = FALSE) +
  # geom_point(data=fgplayer100, aes(x=lastYear, y=pctfg3, colour=Player), size=1, show.legend = FALSE) +
  # geom_line
  xlab('Year') +
  ylab(NULL) +
  ggtitle('3 point success rate by player and year') +
  theme(panel.background=element_rect(fill=NA), panel.grid.major.y=element_line(color="grey", linetype=2),
        plot.title = element_text(hjust = 0.5)) + 
  scale_y_continuous(limits=c(30, 45), breaks=yaxisbreaks, labels=yaxisbreaks) +
  scale_x_continuous(limits=c(1985,2019), breaks=xaxisbreaks)
plotPlayer2000
```

Above graph shows more players are trying 3 point shots than before. even though the average success rate is similar.


Q4. Players who are good at 3-pointers are also good at 2-pointers or free throws?

By regression.

Players who are good at free throws tend to be good at 3-pointers.
However, 2-point field goal success rate is not related with 3-point field goal success rate!!!
Why?

```{r}
linearModel <- lm(pctfg3 ~ pctfg2, data=fgplayer100)
tidy(linearModel)

linearModel2 <- lm(fg3m ~ fgm, data=fgplayer100)
tidy(linearModel2)

linearModel3 <- lm(fg3a ~ fga, data=fgplayer100)
tidy(linearModel3)

linearModel4 <- lm(fg3a ~ fga + fta, data=fgplayer100)
tidy(linearModel4)

linearModel5 <- lm(pctfg3 ~ pctft, data=fgplayer100)
tidy(linearModel5)

linearModel6 <- lm(pctfg2 ~ pctft, data=fgplayer100)
tidy(linearModel6)

linearModel7 <- lm(pctfg3 ~ pctfg2 + pctft, data=fgplayer100)
tidy(linearModel7)
```

When we look at all the players, 2-pointers and 3-pointers are reverse-related. Maybe because of dunk shots?

```{r}
linearModel7 <- lm(pctfg3 ~ pctfg2 + pctft, data=fgplayer)
tidy(linearModel7)
```
Best players (more than 1,000 career 3-point field goals) are good at 2-pointers as well!!!

```{r}
linearModel7 <- lm(pctfg3 ~ pctfg2 + pctft, data=fgplayer1000)
tidy(linearModel7)

linearModel8 <- lm(pctfg3 ~ pctfg2 + pctft, data=fgplayer2000)
tidy(linearModel8)
```

-. Are there any relationship between players’ ages and 3-pointers? Both total and average.

```{r}
fgyearplayer100 <- fgyearplayer %>% filter(Player %in% fgplayer100$Player)
fgyearplayer1000 <- fgyearplayer100 %>% filter(Player %in% fgplayer1000$Player)
fgyearplayer2000 <- fgyearplayer1000 %>% filter(Player %in% fgplayer2000$Player)

xaxisbreaks <- seq(1985, 2019, by=3)
yaxisbreaks <- seq(0, 100, by=5)

plotYearPlayer100 <- ggplot() +
  geom_line(data=fgyearplayer100, aes(x=Year, y=pctfg3, colour=Player), size=1, show.legend = FALSE) +
  geom_line(data=fgyearplayer100, aes(x=Year, y=pctfg2, colour=Player), size=1, linetype="dotted", show.legend = FALSE) +
  xlab('Year') +
  ylab(NULL) +
  ggtitle('3 point shot success rate by player') +
  theme(panel.background=element_rect(fill=NA), panel.grid.major.y=element_line(color="grey", linetype=2),
        plot.title = element_text(hjust = 0.5)) + 
  scale_y_continuous(limits=c(0, 100), breaks=yaxisbreaks, labels=yaxisbreaks) +
  scale_x_continuous(limits=c(1985,2019), breaks=xaxisbreaks)
plotYearPlayer100


xaxisbreaks <- seq(1985, 2019, by=3)
yaxisbreaks <- seq(0, 70, by=5)

plotYearPlayer1000 <- ggplot() +
  geom_line(data=fgyearplayer1000, aes(x=Year, y=pctfg3, colour=Player), size=1, show.legend = FALSE) +
  geom_line(data=fgyearplayer1000, aes(x=Year, y=pctfg2, colour=Player), size=1, linetype="dotted", show.legend = FALSE) +
  xlab('Year') +
  ylab(NULL) +
  ggtitle('3 point shot success rate by player') +
  theme(panel.background=element_rect(fill=NA), panel.grid.major.y=element_line(color="grey", linetype=2),
        plot.title = element_text(hjust = 0.5)) + 
  scale_y_continuous(limits=c(0, 70), breaks=yaxisbreaks, labels=yaxisbreaks) +
  scale_x_continuous(limits=c(1985,2019), breaks=xaxisbreaks)
plotYearPlayer1000

xaxisbreaks <- seq(1985, 2019, by=3)
yaxisbreaks <- seq(25, 55, by=5)

plotYearPlayer2000 <- ggplot() +
  geom_line(data=fgyearplayer2000, aes(x=Year, y=pctfg3, colour=Player), size=1, show.legend = FALSE) +
  geom_line(data=fgyearplayer2000, aes(x=Year, y=pctfg2, colour=Player), size=1, linetype="dotted", show.legend = FALSE) +
  xlab('Year') +
  ylab(NULL) +
  ggtitle('3 point shot success rate by player') +
  theme(panel.background=element_rect(fill=NA), panel.grid.major.y=element_line(color="grey", linetype=2),
        plot.title = element_text(hjust = 0.5)) + 
  scale_y_continuous(limits=c(25, 55), breaks=yaxisbreaks, labels=yaxisbreaks) +
  scale_x_continuous(limits=c(1985,2019), breaks=xaxisbreaks)
plotYearPlayer2000


```

Let's regress.

```{r}
fgyearplayerjoined <- left_join(fgyearplayer, fgplayer, by=c("Player" = "Player"))
fgyearplayerjoined$career = fgyearplayerjoined$Year - fgyearplayerjoined$firstYear + 1

fgyearplayerjoined100 <- fgyearplayerjoined %>% filter(Player %in% fgplayer100$Player)
fgyearplayerjoined1000 <- fgyearplayerjoined100 %>% filter(Player %in% fgplayer1000$Player)
fgyearplayerjoined2000 <- fgyearplayerjoined1000 %>% filter(Player %in% fgplayer2000$Player)

linearModel <- lm(pctfg3.x ~ career, data=fgyearplayerjoined2000)
tidy(linearModel)
linearModel2 <- lm(pctfg3.x ~ career, data=fgyearplayerjoined1000)
tidy(linearModel2)
linearModel3 <- lm(pctfg3.x ~ career, data=fgyearplayerjoined100)
tidy(linearModel3)
linearModel4 <- lm(pctfg3.x ~ career, data=fgyearplayerjoined)
tidy(linearModel4)
```

Really good players are not related with ages/career.
Average players' success rate is increased by 0.4% in one year. Not bad...?

-	Players with high salaries are good at 3-pointers?

2018-2019 season data only
```{r}
nbaInsiderSalaries <- nba_insider_salaries(assume_player_opt_out = T, assume_team_doesnt_exercise = T, return_message = TRUE)

fgplayersalary <- left_join(fgplayer, nbaInsiderSalaries, by=c("Player"="namePlayer"))

fgplayersalary2 <- na.omit(fgplayersalary)
fgplayersalary2$salaryinK = fgplayersalary2$value / 1000
fgplayersalary2$salaryinM = fgplayersalary2$value / 1000000

linearModel <- lm(pctfg3 ~ salaryinM, data=fgplayersalary2)
tidy(linearModel)
linearModel2 <- lm(fg3m ~ salaryinM, data=fgplayersalary2)
tidy(linearModel2)

```
When the salary increases by a million dollar, career success rate of 3-point shots increases by 0.09% only.
It's difficult to say that 3-pointer success rate is the most important factor for one's salary.

-	We would like to explore the importance of three point shooters in a given team by measuring the share of the team’s total salary over time.


-	We want to analyze whether players can drastically improve their three point shooting skills over time or the skill is rather something people are borned with.

There is no dramatic increase in 3-pointer success rate.
Maybe if we can check the players' data from NCAA or high school league, there might be different insight.
However, based on NBA data, no big changes.

-	Show the 3-pointer statistics geographically based on players’ hometowns. Maybe this help illustrates the different basketball playing style across different regions, both domestic and international.

```{r}
playerHometown <- read_csv("PlayerHometown.csv")

fgplayerhometown <- left_join(fgplayer, playerHometown, by=c("Player"="Player"))
fgplayerhometown <- fgplayerhometown %>% filter(not(is.na(State)))
fgplayerhometown <- na.omit(fgplayerhometown)

fgplayerhometownState <- aggregate(fgplayerhometown[, 2:7], list(fgplayerhometown$State), sum)
colnames(fgplayerhometownState)[1] <- "State"
fgplayerhometownState$pctfg3 <- fgplayerhometownState$fg3m / fgplayerhometownState$fg3a * 100
fgplayerhometownState$pctfg2 <- fgplayerhometownState$fgm / fgplayerhometownState$fga * 100
fgplayerhometownState$pctft <- fgplayerhometownState$ftm / fgplayerhometownState$fta * 100

plotState <- ggplot() +
  geom_point(data=fgplayerhometownState, aes(x=State, y=pctfg3, colour=State)) +
  xlab(NULL) +
  ylab(NULL)
plotState

```
